\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-04-27 paper/main.tex
.

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluated on other datasets. Following the results of MolTrans
\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared
it to our ``na\"ive'' predictor as well as \name{} (see Table
\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
pair and the protein split; we were able to reproduce the MolTrans
results (Table \ref{tab:comparison2}), showing a substantial
difference based on the splitting scheme. We additionally computed the
$MicroAUC_p$ score for all considered methods, leading to similar
results as observed on the Yaminishi dataset. As \name{} considers
less drugs than alternative approaches, we execute a one-tailed T-test
with $\alpha=0.01$ over all considered methods. Considering Macro
AUCs, \name{} improves significantly ($p<10^{-4}$, one-sided t-test)
over MolTrans ($t=36.6$), DeepDTI ($t=40.5$), DeepDTA ($t=36.6$), and
DeepConv-DTI ($t=40.5$) on the BioSnap dataset, and significantly
($p<10^{-4}$, one-sided t-test) over DTINet ($t=20.8$) and DTIGEMS+
($t=18.0$) on the Yamanishi dataset; DTI-CDF performs significantly
better than \name{} ($p=0.02$, one-sided t-test). % As
                                % we compare \name{} to 4 different
                                % methods over BioSnap and 3 over
                                % Yamanishi dataset, we add a
                                % Bonferroni correction by multiplying
                                % the $p$-values by 4 and 3,
                                % respectively, leading to $p<0.0002$
                                % ($\alpha=0.01$) for all previous
                                % comparisons except DTI-CDF.
We further perform another one-tailed t-test to compare $MicroAUC_p$
performance. We find that \name{} improved significantly ($p <
10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and
over DTINet ($t=29.6$), DTIGEMS+ ($t=28.0$), and DTI-CDF ($t=9.5$) on
the Yamanishi dataset.

>>>>

As the difference in performance with different splitting schemes is
quite large, we further evaluated additional drug--target interaction
and drug--target affinity prediction methods that were trained and
evaluated on other datasets. Following the results of MolTrans
\cite{MolTrans2020}, we reevaluated DeepDTI \cite{DeepDTI2017},
DeepDTA \cite{DeepDTA2018}, DeepConv-DTI \cite{DeepConvDTI2019}, and
MolTrans itself on the BioSnap dataset \cite{BioSnap2018} and compared
it to our ``na\"ive'' predictor as well as \name{} (see Table
\ref{tab:comparison2}).  MolTrans was evaluated over the drug--target
pair and the protein split; we were able to reproduce the MolTrans
results (Table \ref{tab:comparison2}), showing a substantial
difference based on the splitting scheme. We additionally computed the
$MicroAUC_p$ score for all considered methods, leading to similar
results as observed on the Yaminishi dataset. We test whether \name{}
performs better than the methods we compare against or whether the
observed values for $MicroAUC_p$ and $MacroAUC$ fall within expected
variant; we use a one-tailed T-test with Bonferroni correction for
this test.

Considering Macro AUCs, \name{} improves significantly ($p<10^{-4}$,
one-sided t-test) over MolTrans ($t=36.6$), DeepDTI ($t=40.5$),
DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet
($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
performs significantly better than \name{} ($p=0.02$, one-sided
t-test).  We further perform another one-tailed t-test to compare
$MicroAUC_p$ performance. We find that \name{} improved significantly
($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and
over DTINet ($t=29.6$), DTIGEMS+ ($t=28.0$), and DTI-CDF ($t=9.5$) on
the Yamanishi dataset.
